Joint photometric modeling for improved measurements in imaging surveys.

### 2014-11-18 documents/paper/ms_v0.0.tex
more figs/text

In figure \ref{fig:posteriors}, we show the posterior distributions for 3
stars and 3 galaxies at different magnitudes.

So far we have shown operationally that the XD model does sensible modeling of
the photometry and is more \emph{precise}, but it is important to ask whether
the posteriors are at all \emph{accurate}.  To do so, we rely on comparsions of
the posteriors we get for S82 single epoch data versus the data from the
coadd.  In figure \ref{fig:xx_plots}, we plot the different between the single
epoch and coadd data, as well as our XD posteriors versus the coadd.  We find
that the median of the residuals for our XD posteriors is consistent with
zero, indicating the model is not only precise but also accurate.  We note
this is true for all of the features we model.  Also obvious is the clear
improvement in the photometry over the single epoch data.

In SDSS, star-galaxy classification is accomplished by selecting objects a
small range of PSF minus model magnitudes and calling them stars.  In DR10,
the range is approximately\footnote{The actual algorith involves a combination
of this criterion for $gri$ bands.} $|r_{PSF} - r_{model}| < 0.145$ while the
S82 coadd criterion is $|r_{PSF} - r_{model}| < 0.03$.  In figure \ref{fig:pmm}
we show a normalized histogram of $r$ PSF minus model magnitudes around
$r=21$, for objects that are called stars in the coadd.  Note that our XD
posteriors dramatically tighten the distribution of objects around near zero,
and also indicate the present of interloping galaxies.

>>>>

\begin{figure}
\centering
\includegraphics[clip=true, trim=1.5cm 0.5cm 1.5cm 0.5cm,
  width=16cm]{fig2.png}
\caption{Error ellipses for single epoch S82 photometry (dashed) and XD
posterior distributions for a star (top row) and galaxy (bottom row).  The 
projections are the same as in Figure \ref{fig:contours}.  Ellipses for the XD
posteriors are generally smaller, yet are still consistent with the original 
measurements.
}
\label{fig:posteriors}
\end{figure}

In figure \ref{fig:posteriors}, we show the posterior distributions for a
random star and galaxy in our sample.  We find generally that projections of 
the posterior are consistent with the original measurements, but are of higher 
precision.  Figure \ref{fig:error_rates} quantifies this, showing the median 
uncertainty as a function of $r$ magnitude.  We find that the median
uncertainties of our XD posteriors are a significant improvement to the
catalog data.  For instance, at $r\sim20$, the uncertainty in the DR10 $r$ PSF 
magnitude is reduced from 0.28 to 0.2 \rf{DOUBLE CHECK} and more dramatically
the error in the $u-g$ model color is reduced from 0.5 to 0.3.  It is not
surprising that the improvements are most significant where the data is more 
noisy since this is where an informative prior can make the most difference.

\begin{figure}
\centering
\includegraphics[clip=true, trim=1.5cm 0.5cm 1.5cm 0.5cm,
  width=16cm]{fig3.png}
\caption{Plotted are median values for the 1 sigma uncertainty in $r$ PSF
magnitude, $r$
psf minus model magnitude, and the four model colors used.  The solid black 
curve is for the Stripe 82 coadd, while the dotted black and red curves are for
single epoch Stripe 82 and DR10 data, respectively.  The dashed curves show the
median uncertainty for the XD posteriors.  While not quite as precise as the 
coadd data, the XD posteriors are a significant improvement over the catalog
data.
}
\label{fig:error_rates}
\end{figure}

So far we have shown operationally that the XD model does sensible modeling of
the photometry and is more \emph{precise}, but it is important to ask whether
the posteriors are at all \emph{accurate}.  To do so, we rely on comparsions of
the posteriors we get for S82 single epoch data versus the data from the
coadd.  In figure \ref{fig:xx_plots}, we plot the single
epoch data against coadd measurements, as well as our XD posteriors versus the
coadd.  We find that the XD posteriors are in good agreement with the coadd
data -- there is no obvious biases or deviations appart from scatter due to
noise.  We note that there is a slight offset to lower values of $r$ PSF
magnitudes at the faintest end ($r \gtrsim 21.5$).  This is an artifact due to
the fact that none of the components we used were constrained to have mean
around or greater than $r \sim 21.5$.  As a result, the posteriors will
naturally pull the magnitude values down.  A simple fix is to force a few
components of the XD mixture to have means at faint magnitude values, but given
the size of this effect we have not done so for the purposes here.  Note a
similar effect will also occur at the bright end of our sample, but is less 
significant since the measurement uncertainties are lower.  As a additional 
demonstration of the reasonability of our method, we show the color-magnitude
diagram for DR10 data of the globular cluster Messier 3 in Figure
\ref{fig:m3}.  Compared to the DR10 data the XD posterior is clearly much
cleaner at faint $r$ magnitudes, and is more concentrated across much of the
main sequence.  Note that the width of the main sequence around $r \sim 19$ is
not dramatically changed from DR10.  This is because the measurement
uncertainties are fairly small so the DR10 value is the dominant contributor
to the posterior.  In other words, the XD values seem to be respectful of the 
data and do not seem to overly `denoise'.

\begin{figure}
\centering
\includegraphics[clip=true, trim=1.5cm 0.5cm 1.5cm 0.5cm,
  width=16cm]{fig4.png}
\caption{In the top row, we plot 2D histograms of $r$ PSF magnitudes, $u-g$
model color, and $i-z$ model color for single epoch Stripe 82 data versus the
corresponding coadd values.  In the bottow row, we plot the same for a single
sample from the XD posteriors of the data in the top.  Generally speaking the
XD posteriors are very consistent with the coadd photometry, indicating that 
the models are indeed acurate.
}
\label{fig:xx_plots}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=16cm]{fig5.png}
\caption{The $g-r$ versus $r$ color-magnitude diagram from the outskirts of
the globular cluster Messier 3.  The left panel shows the original DR10
measurements and the right shows the median value from the XD posteriors.  Note
the main sequence is generally more concentrated, especially at faint
magnitudes.
}
\label{fig:m3}
\end{figure}

In SDSS, star-galaxy classification is accomplished by selecting objects a
small range of PSF minus model magnitudes and calling them stars.  In DR10,
the range is approximately\footnote{The actual algorithm involves a combination
of this criterion for $gri$ bands.} $|r_{PSF} - r_{model}| < 0.145$ while the
S82 coadd criterion is $|r_{PSF} - r_{model}| < 0.03$.  In figure \ref{fig:pmm}
we show a histogram of $r$ PSF minus model magnitudes around $r=19.5$
20.5, and 21.5 for objects that are called `stars' in the coadd.  Note that
our XD posteriors dramatically tighten the distribution of objects to values
very close to zero.  This implies that the quality of the star-galaxy
classification could be substantially improved by selecting on values produced
by the XD posteriors.

\begin{figure}
\centering
\includegraphics[width=16cm]{fig6.png}
\caption{Histograms of the values of the $r$ PSF minus model magnitudes for 
objects within 0.15 mag of $r = 19.5$, 20.5, and 21.5 that were labeled as 
`stars' in the Stripe 82 coadd.  The dotted histograms shows the distribution
for single epoch Stripe 82 data, while the solid histograms show those for the
XD posteriors.  The XD values are strongly peaked near zero, while the single
epoch values are more distributed, suggesting star-galaxy classification
could be improved by relying on the XD values.  
}
\label{fig:pmm}
\end{figure}
